---
title: "Batch effects adjustments"
author: "Sharonhe"
date: "November 8, 2016"
output: html_document
---

## Dependencies

```{r}
library(devtools)
library(Biobase)
library(sva)
library(bladderbatch)
library(snpStats)
```

## Laod the data

Data are gene expression measurements from a bladder cancer study: [Gene expression in the urinary bladder: a common carcinoma in situ gene expression signature exists disregarding histopathological classification.](http://cancerres.aacrjournals.org/content/64/11/4040.full.pdf) The data can be loaded from the [bladderbatch](http://bioconductor.org/packages/release/data/experiment/html/bladderbatch.html) data package. 

```{r}
data(bladderdata)
pheno = pData(bladderEset)
edata = exprs(bladderEset)
```

## Adjustment for batch effects with a linear model

```{r}
mod = model.matrix(~as.factor(cancer) + as.factor(batch),data=pheno)
fit = lm.fit(mod,t(edata))
hist(fit$coefficients[2,],col=2,breaks=100)

# This will only work if the batch effects aren't too highly correlated with the outcome.

table(pheno$cancer,pheno$batch)
```

## Adjusting for batch effects with Combat

Combat returns a "cleaned" data matrix after batch effects have been removed.

https://www.bioconductor.org/packages/release/bioc/html/sva.html

```{r}
batch = pheno$batch
modcombat = model.matrix(~1, data=pheno) # any other adjustment we want to do
modcancer = model.matrix(~cancer, data=pheno) # actual association we want to test
combat_edata = ComBat(dat=edata, batch=batch, mod=modcombat, par.prior=TRUE, prior.plots=FALSE)
combat_fit = lm.fit(modcancer,t(combat_edata)) # asscociation after removing batch effect
hist(combat_fit$coefficients[2,],col=2,breaks=100)
```

## Adjusting for batch effects with sva (surrogate variable analysis)

First we need to estimate the surrogate variables. To do this, we need to build a model with any known adjustment variables and the variable we care about `mod` and another model with only the adjustment variables. Here we won't adjust for anything to see if `sva` can "discover" the batch effect. 

```{r}
mod = model.matrix(~cancer,data=pheno) # the variable we care about
mod0 = model.matrix(~1, data=pheno) # unknown variable
sva1 = sva(edata,mod,mod0,n.sv=2) # number of expected surrogate variables == 2
names(sva1)
dim(sva1$sv)
```

See if any of the variables correlate with batch

```{r}
summary(lm(sva1$sv ~ pheno$batch)) # exam whether any sv is highly correlated with batch
# results showed that Response Y2 is highly associated with batch

boxplot(sva1$sv[,2] ~ pheno$batch)
points(sva1$sv[,2] ~ jitter(as.numeric(pheno$batch)),col=as.numeric(pheno$batch))
```

Add the surrogate variables to the model matrix and perform the model fit

```{r}
modsv = cbind(mod,sva1$sv)
fitsv = lm.fit(modsv,t(edata))
hist(fitsv$coefficients[2,], col=2, breaks=100)
```


## Session information

```{r session_info}
devtools::session_info()
```

This document was processed on: `r Sys.Date()`.
