---
title: "Batch effects adjustments"
author: "Sharonhe"
date: "November 8, 2016"
output: html_document
---

## Dependencies

```{r}
library(devtools)
library(Biobase)
library(sva)
library(bladderbatch)
library(snpStats)
```

## Laod the data

Data are gene expression measurements from a bladder cancer study: [Gene expression in the urinary bladder: a common carcinoma in situ gene expression signature exists disregarding histopathological classification.](http://cancerres.aacrjournals.org/content/64/11/4040.full.pdf) The data can be loaded from the [bladderbatch](http://bioconductor.org/packages/release/data/experiment/html/bladderbatch.html) data package. 

```{r}
data(bladderdata)
pheno = pData(bladderEset)
edata = exprs(bladderEset)
```

## Adjustment for batch effects with a linear model

```{r}
mod = model.matrix(~as.factor(cancer) + as.factor(batch),data=pheno)
fit = lm.fit(mod,t(edata))
hist(fit$coefficients[2,],col=2,breaks=100)
```


## Session information

```{r session_info}
devtools::session_info()
```

This document was processed on: `r Sys.Date()`.
