---
title: "Dimension Reduction svd analysis"
author: "Sharonhe"
date: "November 7, 2016"
output: html_document
---

The purpose of dimention reduction is to find patterns in matrix of data. 

## Dependencies

```{r}
library(devtools)
library(Biobase)
```

## Load data

```{r}
con =url("http://bowtie-bio.sourceforge.net/recount/ExpressionSets/montpick_eset.RData")
load(file=con)
close(con)
mp = montpick.eset
pdata=pData(mp)
edata=as.data.frame(exprs(mp))
fdata = fData(mp)
ls()
```

## Calculate the singular vectors

```{r}
edata = edata[rowMeans(edata) > 100, ] # filter the data
dim(edata)
edata = log2(edata + 1) # transform the data
edata_centered = edata - rowMeans(edata) # center the data, this is important for the following svd analysis
svd1 = svd(edata_centered) 
names(svd1)
```

## Look at the percent variance explained

```{r}
plot(svd1$d,ylab="Singular value",col=2)
plot(svd1$d^2/sum(svd1$d^2),ylab="Percent Variance Explained",col=2)
```

## Plot top two principal components

```{r}
par(mfrow=c(1,2))
plot(svd1$v[,1],col=2,ylab="1st PC")
plot(svd1$v[,2],col=2,ylab="2nd PC")
```

## Plot PC1 vs. PC2

Plot PC1 versus PC2 to see if we can see any "clusters" or "groups".

```{r}
plot(svd1$v[,1],svd1$v[,2],col=2,ylab="2nd PC",xlab="1st PC")
```

Color by different variables to see if clusters stand out. 

```{r}
plot(svd1$v[,1],svd1$v[,2],ylab="2nd PC",
     xlab="1st PC",col=as.numeric(pdata$study))
```

Make boxplots comparing the PC for different levels of known covariates.

```{r}
boxplot(svd1$v[,1] ~ pdata$study,border=c(1,2))
points(svd1$v[,1] ~ jitter(as.numeric(pdata$study)),col=as.numeric(pdata$study))
```

## PCs versus SVs

```{r}
pc1 = prcomp(edata) # This is not exactly the principal components
plot(pc1$rotation[,1],svd1$v[,1])
```

To get the actual PCs we have to subtract the column means rather than the row means when normalizing. 

```{r}
edata_centered2 = t(t(edata) - colMeans(edata))
svd2 = svd(edata_centered2)
plot(pc1$rotation[,1],svd2$v[,1],col=2)
```

## Pay attention to outliers

```{r}
edata_outlier = edata_centered
edata_outlier[6,] = edata_centered[6,] * 10000
svd3 = svd(edata_outlier)
plot(svd1$v[,1],svd3$v[,1],xlab="Without outlier",ylab="With outlier")
plot(svd3$v[,1],edata_outlier[6,],col=4) # It turns out the new top singular vector is perfectly correlated with the outlying gene, so we need to pay attention to outliers.
```

## Session information

Here is the session information 

```{r session_info}
devtools::session_info()
```

It is also useful to compile the time the document was processed. This document was processed on: `r Sys.Date()`.