---
title: "Generalized linear model: Logistic regression"
author: "Sharonhe"
date: "November 9, 2016"
output: html_document
---

## Dependencies

```{r}
library(devtools)
library(Biobase)
library(snpStats)
library(broom)
library(MASS)
library(DESeq2)
```

## Load the data

Sample SNP data are from a case-control genome-wide association study. 

```{r}
data(for.exercise)
use <- seq(1, ncol(snps.10), 10)
sub.10 <- snps.10[,use]
```

## Single logistic regression

The coefficient is the change in log-odds for a one unit decrease (because homozygous major allele is coded 1) in the number of copies of the minor allele.

```{r}
snpdata = sub.10@.Data
status = subject.support$cc
snp1 = as.numeric(snpdata[,1]) # analyze the 1st snp
table(snp1) # check for missing values: in results, 0 refers to missing value
snp1[snp1==0] = NA
glm1 = glm(status ~ snp1,family="binomial")
tidy(glm1)
```

## If we need to code a dominant model

Only an association of risk with the two copies of the common allele, now the coefficient on `snp1_dom` is the increase in log odds associated with two copies of the major allele. 

```{r}
snp1_dom = (snp1 == 1)
glm1_dom = glm(status ~ snp1_dom,family="binomial")
tidy(glm1_dom)
tidy(glm1)
```

## Adjust logistic regression on principle components

```{r}
# Compute the pcs
xxmat <- xxt(sub.10, correct.for.missing=FALSE)
evv <- eigen(xxmat, symmetric=TRUE)
pcs <- evv$vectors[,1:5]
# adjust for pcs
glm2 = glm(status ~ snp1 + pcs[,1:5],family="binomial")
tidy(glm2)
```

## Fit many glms at once

The `snps.rhs.tests` function computes an asymptotic chi-squared statistic.

```{r}
# fit many glms
glm_all = snp.rhs.tests(status ~ 1,snp.data=sub.10) # ~1 refers to no adjustment for the model
slotNames(glm_all)
qq.chisq(chi.squared(glm_all),df=1)
# adjust for pcs
glm_all_adj = snp.rhs.tests(status ~ pcs,snp.data=sub.10)
qq.chisq(chi.squared(glm_all_adj),df=1)
```

## Session information

```{r session_info}
devtools::session_info()
```

This document was processed on: `r Sys.Date()`.
