---
title: "Many regressions"
author: "Sharonhe"
date: "November 8, 2016"
output: html_document
---

## Dependencies

```{r}
library(devtools)
library(Biobase)
library(limma)
library(edge)
```

# Load the data

Data from the paper [Detection of redundant fusion transcripts as biomarkers or disease-specific therapeutic targets in breast cancer.](http://www.ncbi.nlm.nih.gov/pubmed/22496456) that uses data from different normal human tissues

```{r}
con =url("http://bowtie-bio.sourceforge.net/recount/ExpressionSets/bottomly_eset.RData")
load(file=con)
close(con)
bot = bottomly.eset
pdata=pData(bot)
edata=as.matrix(exprs(bot))
fdata = fData(bot)
```

## Transform and filter the data

```{r}
edata = log2(as.matrix(edata) + 1)
edata = edata[rowMeans(edata) > 10, ]
```

## Fit many regression models at once. 

```{r}
mod = model.matrix(~ pdata$strain)
fit = lm.fit(mod,t(edata))
names(fit)

# Compare to output of `lm`
fit$coefficients[,1]
tidy(lm(as.numeric(edata[1, ]) ~ pdata$strain))

# Check the coefficients across genes
par(mfrow=c(1,2))
hist(fit$coefficients[1,],breaks=100,col=2,xlab="Intercept")
hist(fit$coefficients[2,],breaks=100,col=2,xlab="Strain")
abline(v=0,lwd=3,col=1)

# Fit many regressions with an adjustment
mod_adj = model.matrix(~ pdata$strain + as.factor(pdata$lane.number))
fit_adj = lm.fit(mod_adj,t(edata))
fit_adj$coefficients[,1]
```

## Fit many regressions with the limma package

```{r}
mod_adj = model.matrix(~ pdata$strain + as.factor(pdata$lane.number))
fit_limma = lmFit(edata,mod_adj)
names(fit_limma)
fit_limma$coefficients[1,]
fit_adj$coefficients[,1] # results are same as fit_limma$coefficients[1,]
```

## Fit many regressions with the edge package

```{r}
edge_study = build_study(data=edata,grp=pdata$strain,adj.var=as.factor(pdata$lane.number))
fit_edge = fit_models(edge_study)
summary(fit_edge)
fit_edge@beta.coef[1,] 
fit_limma$coefficients[1,] # results are same as fit_edge@beta.coef[1,] 
```

## Session information

```{r session_info}
devtools::session_info()
```

This document was processed on: `r Sys.Date()`.