---
title: "Permutation"
author: "Sharonhe"
date: "November 11, 2016"
output: html_document
---

## Dependencies

```{r}
library(devtools)
library(Biobase)
library(limma)
library(edge)
library(genefilter)
```

# Load the data

Data are from the paper [Evaluating gene expression in C57BL/6J and DBA/2J mouse striatum using RNA-Seq and microarrays.](http://www.ncbi.nlm.nih.gov/pubmed?term=21455293) that is a comparative RNA-seq analysis of different mouse strains.

```{r}
con =url("http://bowtie-bio.sourceforge.net/recount/ExpressionSets/bottomly_eset.RData")
load(file=con)
close(con)
bot = bottomly.eset
pdata=pData(bot)
edata=as.matrix(exprs(bot))
fdata = fData(bot)
edata = log2(as.matrix(edata) + 1)
edata = edata[rowMeans(edata) > 10, ]
```

## Calculate t- or F-statistics

```{r}
tstats_obj = rowttests(edata,pdata$strain)
hist(tstats_obj$statistic,col=2,xlim=c(-5,2))
```

## Permute the sample labels

We can now permute the sample labels using the `sample` function in R. 

```{r}
set.seed(135)
strain = pdata$strain
strain0 = sample(strain)
tstats_obj0 = rowttests(edata,strain0)
hist(tstats_obj0$statistic,col=2,xlim=c(-5,2))
```

## Compare the observed and permuted statistics

```{r}
quantile(tstats_obj0$statistic)
quantile(tstats_obj$statistic)
```

## Session information

```{r session_info}
devtools::session_info()
```
      
This document was processed on: `r Sys.Date()`.

